### 2024-02-09 doc/15yrsChange.Rmd
Update Outline and Add Relevant Literature Notes

Non-endemic pests and pathogens facilitate invasin of non-endemic plants [@guo_impacts_2023 and refs therein].

@@

>>>>

Non-endemic pests and pathogens facilitate invasion of non-endemic plants [@guo_impacts_2023 and refs therein].

@@

Recruitment Failure Paragraph Outline

- Over-abundant herbivores can have a profound impact on carbon cycling in temperate deciduous forests
- There is a documented overabundance of white-tailed deer (*Odocoileus virginianus*) (*Publication referenced in McGravey et al. 2013*) which can be attributed to human influence (*Publication referenced in McGravey et al 2013*)
- White-tailed deer consume seedlings and saplings, negatively impacting seedling and sapling survival, density and growth (*Publications referenced in McGravey et al. 2013*, *Holm et al. 2013*)
- This can lead to decreased understory diversity, species richness and decreased abundance of dominant species in a typical forest (in this case Quercus spp.) (*Holm et al. 2013*)
- Define or contextualize non-endemic plants (Look for a reference)
  - Working definition: Plants that were not historically present in an area, region or ecosystem or were found at different abundances or densities
- Deer selectively browse on palatable species, and some non-endemic species, including pawpaw, are considered non-palatable by deer (*McGravey et al. 2013*)  This can create dense stands of non-endemic species (*Knauer et al. 2023*)
- Non-endemic species may be able to utilize niche space in a forest faster or more efficiently (Reference - considered common knowledge?)
- As a result, they may be able to out compete endemic species
- These nuisance species can contribute to recruitment failure
  - Consider defining and contextualizing recruitment failure a little more here
  - The effects of these recruitment failure on species composition and forest structure are often not apparent for decades (*McGravey et al. 2013* - potentially move to regeneration debt paragraph).
